---
title: "Family Planning Shiny App"
author: "Marcela Guerra Osorno"
date: "May 1, 2018"
output: html_notebook
runtime: shiny
---

```{r Library, include=FALSE}
library(shiny)
library(rworldmap)
library(RColorBrewer)
library(classInt)
library(shinyjs)
```
#INTRODUCTION
### The United Nations Foundation links the UN's work with others around the world, mobilizing the energy and expertise of business and non-governmental organizations to help the UN tacke issues including climate change, global health, peace and security, women's empowerment, poverty eradication, energy access, and U.S.-UN relations.
### On the other hand, Family Planning 2020 (FP2020) is a global partnership that supports the rights of women and girls to decide, freely, and for themselves, whether, when, and how many children they want to have.

#ABOUT THE SHINY APP
### The following code runs a Shiny App designed to visualize data on family planning, by using heat maps, index plots, summary statistics, and a simple linear regression model.
### The dataset contains information on 69 observations (based on the 69 focus countries where FP2020 currently works), and three variables: 'unmet need for modern contraception', 'adolescent birth rate','fertility rate' 
```{r Panels}
ui <- fluidPage(
  tags$head(
    tags$style(HTML("
                    @import url('//fonts.googleapis.com/css?family=Lobster|Cabin:400,700');
                    "))
    ),
  titlePanel(h1("Global Data on Family Planning", 
                style = "font-family:'Lobster',cursive;
                font-weight: 500; line-height: 1.1; 
                color: #4d3a7d;")),
  
  # Side panel
  sidebarLayout(
    sidebarPanel(
      actionButton("do", "About FP2020"),
      verbatimTextOutput("text"),
      tags$head(tags$style("#text{font-size: 14px
                           
                           }"
                         )
      ),
      radioButtons("plot", "Choose your plot",
                   c("Scatterplot Unmet need", "Map Unmet need", "Map Adolescent Birth Rate", 
                     "Map Fertility Rate", "Fertility Rate ~ Unmet need")),
      "Created by Marcela Guerra Osorno"
      ),
    
    # Main panel
    mainPanel(
      tabsetPanel(
        tabPanel("Plots", (plotOutput("plotdot"))),
        tabPanel("Summary", "Summary statistics for the percentage of women with an unmet need for 
                 modern contraception:", (verbatimTextOutput("statsUnmet")), 
                 "Summary statistics for the number of births to adolescent females, aged 15-19 per 1,000:", 
                 (verbatimTextOutput("statsABR")),
                 "Summary statistics for the fertility rate per woman:", 
                 (verbatimTextOutput("statsFR")),
                 "Summary statistics for a Simple Linear Regression Model: Fertility Rate ~ Unmet need",
                 (verbatimTextOutput("statsFP"))),
        tabPanel("List of Countries", ("In the following link you can consult the complete list of the 69 
                                       focus countries where FP2020 is currently working"), (uiOutput("link")) )
      )
    )
  ))
```

```{r inputs and outputs}
server <- function(input, output) {
  
  observeEvent(input$do, {
    output$text <- renderText({"Family Planning 2020 (FP2020) uses a suite of quantitative metrics to monitor 
      progress annually across the 69 focus countries. The following dataset corresponds to the the percentage 
      of fecund women of reproductive age who want no more children or to postpone having the next child, but 
      are not using a contraceptive method, plus women who are currently using a traditional method of family 
      planning. Women using a traditional method are assumed to have an unmet need for modern contraception."
    })
  })
  url <- a("List of 69 focus countries", 
           href="http://progress.familyplanning2020.org/en/appendices/fp2020-focus-countries")
  output$link <- renderUI({
    tagList("URL:", url)
  })
  # Import data
  FP2020<-read.csv("./Data/FP2020.csv", header = TRUE)
  attach(FP2020)
  
  output$plotdot <- renderPlot({
    
    if (input$plot == "Scatterplot Unmet need") {
      title <- "Percentage of women with an unmet need for modern contraception" 
      plot(Unmet, main = title, col='orchid')
      text(Unmet, labels = Country, cex = .7, pos = 2, col = 'navy')
    }
    else if (input$plot == "Map Unmet need") {
      n <- joinCountryData2Map(FP2020, joinCode="NAME", nameJoinColumn="Country")
      classInt <- classInt::classIntervals( n[["Unmet"]], n=4, style="jenks")
      catMethod = classInt[["brks"]]
      colourPalette <- RColorBrewer::brewer.pal(4,'RdPu')
      mapParams<- mapCountryData(n, nameColumnToPlot="Unmet", addLegend=FALSE, mapTitle="Percentage of women with an unmet need for modern contraception",
                                 catMethod=catMethod, 
                                 colourPalette=colourPalette )
      do.call( addMapLegend, c( mapParams, legendLabels="all", legendWidth=0.5, legendIntervals="data", 
                                legendMar = 2 ) )  
    }
    
    else if (input$plot == "Map Adolescent Birth Rate") {
      n <- joinCountryData2Map(FP2020, joinCode="NAME", nameJoinColumn="Country")
      classInt <- classInt::classIntervals( n[["ABR"]], n=4, style="jenks")
      catMethod = classInt[["brks"]]
      colourPalette <- RColorBrewer::brewer.pal(4,'Purples')
      mapParams<- mapCountryData(n, nameColumnToPlot="ABR", addLegend=FALSE, mapTitle="Number of births to adolescent females, aged 15-19 per 1,000 adolescent females",
                                 catMethod=catMethod, 
                                 colourPalette=colourPalette )
      do.call( addMapLegend, c( mapParams, legendLabels="all", legendWidth=0.5, legendIntervals="data", 
                                legendMar = 2 ))
      
    }
    
    else if (input$plot == "Map Fertility Rate") {
      n <- joinCountryData2Map(FP2020, joinCode="NAME", nameJoinColumn="Country")
      classInt <- classInt::classIntervals( n[["FR"]], n=4, style="jenks")
      catMethod = classInt[["brks"]]
      colourPalette <- RColorBrewer::brewer.pal(4,'BuGn')
      mapParams<- mapCountryData(n, nameColumnToPlot="FR", addLegend=FALSE, mapTitle="Fertility Rate: Number of children per woman",
                                 catMethod=catMethod, 
                                 colourPalette=colourPalette )
      do.call( addMapLegend, c( mapParams, legendLabels="all", legendWidth=0.5, legendIntervals="data", 
                                legendMar = 2 ) )  
    }
    
    else if (input$plot == "Fertility Rate ~ Unmet need") {
      title <- "Fertility Rate vs. Unmet need" 
      plot(FR~Unmet, main = title, col='orchid')
      text(y= "Fertility Rate", x="Unmet", labels = Country, cex = .7, pos = 2, col = 'navy')
    }
    
  })
  
  output$statsUnmet <- renderPrint({
    summary(FP2020$Unmet)
  })
  
  output$statsABR <- renderPrint({
    summary(FP2020$ABR)
  })
  
  output$statsFR <- renderPrint({
    summary(FP2020$FR)
  })
  
  output$statsFP <- renderPrint({
    FPreg<-lm(FP2020$FR~FP2020$Unmet)
    summary(FPreg)
  })
  
}
```

### The heat map allows us to visualize which countries have a higher unmet need percentage, higher adolescent birth rate,  and higher fertility rate

### Frome the simple regression model 'Fertility Rate' ~ 'Unmet Need', we can see that the 'Unmet Need' is significant for explaining 'Fertility Rate'. It makes sense because at a higher level of unmet need for contraception, we would expect a higher fertility rate.

```{r run application}
shinyApp(ui = ui, server = server)
```